---
title: "Bitcoin price prediction"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading the required libraries

```{r}
library(forecast)
library(tseries)
require(graphics)
library(ggplot2)
```

Load the file into R

```{r}
load("rdas/rawdata.rda") 
```
```{r}
nrow(rawdata)
```

## Plot timeseries

Create the model using the first 71 rows. Then test the model on the remaining 6 rows later

```{r}
total_timeser <- ts(rawdata$Price)
indata <- rawdata[1:29,]
timeser <- ts(indata$Price)
plot(timeser)
```

## Smoothing the series - Moving Average Smoothing

```{r}
w <-1
smoothedseries <- filter(timeser, 
                         filter=rep(1/(2*w+1),(2*w+1)), 
                         method='convolution', sides=2)

```

width= window, method convolution is moving average, sides=2, is a two sided filter

Smoothing left end of the time series

```{r}
diff <- smoothedseries[w+2] - smoothedseries[w+1]
for (i in seq(w,1,-1)) {
  smoothedseries[i] <- smoothedseries[i+1] - diff
}

```

Smoothing right end of the time series

```{r}
n <- length(timeser)
diff <- smoothedseries[n-w] - smoothedseries[n-w-1]
for (i in seq(n-w+1, n)) {
  smoothedseries[i] <- smoothedseries[i-1] + diff
}
```

## Plot the smoothed time series

```{r}
timevals_in <- indata$Months
plot(timeser)
lines(smoothedseries, col="blue", lwd=2)
```

## Model building

Building a model on the smoothed time series using classical decomposition

First convert the time series to a dataframe

```{r}
smootheddf <- as.data.frame(cbind(timevals_in, as.vector(smoothedseries)))
colnames(smootheddf) <- c('Months', 'Price')
```

Fitting a multiplicative model with trend and seasonality to the data

Seasonality will be modeled using a sinusoid function

```{r}
lmfit <- lm(Price ~ sin(0.5*Months) * poly(Months,3) + cos(0.5*Months) * poly(Months,3)
            + Months, data=smootheddf)
global_pred <- predict(lmfit, Month=timevals_in)
summary(global_pred)
plot(timeser)
lines(smoothedseries, col="blue", lwd=2)
lines(timevals_in, global_pred, col='red', lwd=2)
```

## ARMA series

Extract locally predictable series and model it as an ARMA series

```{r}
local_pred <- timeser-global_pred
plot(local_pred, col='red', type = "l")
```
 
ACF plot 
 
```{r}
acf(local_pred)
```

PACF plot

```{r}
acf(local_pred, type="partial")
```

## Armafit

```{r}
armafit <- auto.arima(local_pred)
```

Plot armafit

```{r}
tsdiag(armafit)
```

```{r}
armafit
```

## Check if the residual series is white noise

```{r}
resi <- local_pred-fitted(armafit)
```

Augmented Dickey-Fuller Test

```{r}
adf.test(resi,alternative = "stationary")
```

KPSS Test for Level Stationarity

```{r}
kpss.test(resi)
```

# Model evaluation using MAPE

Make a prediction for the last 6 months

```{r}
outdata <- rawdata[30:32,]
timevals_out <- outdata$Months

global_pred_out <- predict(lmfit,data.frame(Months =timevals_out))

fcast <- global_pred_out
```

Compare our prediction with the actual values, using MAPE

```{r}
MAPE_class_dec <- accuracy(fcast,outdata[,2])[5]
MAPE_class_dec
```

Plot the predictions along with original values, to get a visual feel of the fit

```{r}
class_dec_pred <- c(ts(global_pred),ts(global_pred_out))
plot(total_timeser, col = "black")
lines(class_dec_pred, col = "red")
```

So, that was classical decomposition, now let's do an ARIMA fit

## ARIMA fit
Autoarima

```{r}
autoarima <- auto.arima(timeser)
autoarima
```

Plot autoarima

```{r}
tsdiag(autoarima)
```
 
Plot autoarima fit
 
```{r}
plot(autoarima$x, col="black")
lines(fitted(autoarima), col="red")
```

Check if the residual series is white noise

```{r}
resi_auto_arima <- timeser - fitted(autoarima)
```

Augmented Dickey-Fuller Test

```{r}
adf.test(resi_auto_arima,alternative = "stationary")
```

KPSS Test for Level Stationarity

```{r}
kpss.test(resi_auto_arima)
```

## Model evaluation using MAPE

```{r}
fcast_auto_arima <- predict(autoarima, n.ahead = 3)

MAPE_auto_arima <- accuracy(fcast_auto_arima$pred,outdata[,2])[5]
MAPE_auto_arima
```

## Plot the predictions along with original values, to get a visual feel of the fit

```{r}
auto_arima_pred <- c(fitted(autoarima),ts(fcast_auto_arima$pred))
plot(total_timeser, col = "black")
lines(auto_arima_pred, col = "red")
lines(fcast_auto_arima$pred, col = "blue")
```